---
title: "Data visualization"
---



```{r setup, include=FALSE}
# set knit options
knitr::opts_chunk$set(
  fig.width=9, 
  fig.height=5, 
  fig.retina=3,
  fig.align="center",
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

# source in functions
source(here::here("R", "funcs.R"))

```




# In-class example


Here's the code we'll be using in class. Download it and store it with the rest of your materials for this course. If simply clicking doesn't trigger download, you should right-click and select "save link as...". 

* Day one: {{< fas fa-code >}} [Five graphs](/files/example-scripts/five-graphs.R)



## Presidential elections code-through


We're working with `elections_historic` from the `socviz` package:

```{r}
library(socviz)
elections_historic
```


We start with `ggplot()`:

```{r}
ggplot()
```


Tell it what data we are plotting:

```{r}
ggplot(elections_historic)
```


Map the x and y-axis within `aes()`:

```{r}
ggplot(elections_historic, aes(x = popular_pct, y = ec_pct))
```


Add a point geometry:


```{r}
ggplot(elections_historic, aes(x = popular_pct, y = ec_pct)) + geom_point()
```


Go back and map color and shape aesthetics:


```{r}
ggplot(elections_historic, aes(x = popular_pct, y = ec_pct,
                               color = win_party, shape = two_term)) + geom_point()
```


Add labels by mapping the label aesthetic and adding a `geom_label` geometry:


```{r}
ggplot(elections_historic, aes(x = popular_pct, y = ec_pct,
                               color = win_party, shape = two_term, 
                               label = winner_label)) + 
  geom_point() +  
  geom_text() 
```



### Bells and whistles


Here's some other stuff I'm adding to make this graph better:


```{r}
library(ggrepel)

ggplot(elections_historic, aes(x = popular_pct, y = ec_pct,
                               color = win_party, shape = two_term, 
                               label = winner_label)) + 
  geom_point() +  
  geom_text_repel() + 
  labs(x = "Percent of popular vote", 
       y = "Percent of Electoral College vote", 
       title = "Presidential Elections: Popular & Electoral College Margins",
       subtitle = "1824-2016", color = NULL, size = NULL) + 
  scale_y_continuous(labels = scales::percent) + 
  scale_x_continuous(labels = scales::percent) + 
  scale_color_manual(values = c(yellow, red, blue, "gray")) + 
  geom_hline(yintercept = 0.5, size = 1.4, color = "gray80") +
    geom_vline(xintercept = 0.5, size = 1.4, color = "gray80") +  theme(legend.position = "none") 
```



## Recessions


The recession graph from class:


```{r, echo. = TRUE}
ggplot(data = economics, aes(x = date, y = uempmed)) + 
  geom_line()
```



## Organs


The organ data plot:


```{r, echo = TRUE}
ggplot(data = organdata, aes(x = donors, fill = opt)) + 
  geom_histogram()
```

