---
title: "DAGs"
---



```{r setup, include=FALSE}
# set knit options
knitr::opts_chunk$set(
  fig.width=9, 
  fig.height=5, 
  fig.retina=3,
  fig.align="center",
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

# source in functions
source(here::here("R", "funcs.R"))

```





## In-class example


Here's the code we'll be using in class. Download it and store it with the rest of your materials for this course. If simply clicking doesn't trigger download, you should right-click and select "save link as..."

- [<i class="fas fa-file-archive"></i> `08-dags.R`](/files/example-scripts/08-dags.R)




## Example


We write dags using `ggdag` package:


```{r}
library(ggdag)
library(tidyverse)
```



Follows the same syntax as `lm()`, but using `dagify()`:


```{r}
dag = dagify(Y ~ X + A + B,
       A ~ B + C,
       X ~ B + C + D,
       C ~ D,
       exposure = "X", 
       outcome = "Y")
```


Plot:


```{r}
ggdag(dag)
```



Look at all the paths:


```{r}
ggdag_paths(dag)
```


Identify controls:


```{r}
ggdag_adjustment_set(dag)
```

