---
title: "Uncertainty"
---



```{r setup, include=FALSE}
# set knit options
knitr::opts_chunk$set(
  fig.width=9, 
  fig.height=5, 
  fig.retina=3,
  fig.align="center",
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

# source in functions
source(here::here("R", "funcs.R"))

```




## In-class example


Here's the code we'll be using in class. Download it and store it with the rest of your materials for this course. If simply clicking doesn't trigger download, you should right-click and select "save link as..."

- [<i class="fas fa-file-archive"></i> `09-uncertainty.R`](/files/example-scripts/09-uncertainty.R)



## Show sample estimates vary


Load the data:

```{r}
# libraries
library(tidyverse)
library(stevedata)
library(moderndive)

# data
gss_abortion
```


What's the true population parameter? Pretending dataset is full population:

```{r}
## abortion true
gss_abortion |> 
  summarise(avg_abnomore = mean(abnomore, na.rm = TRUE))
```


Show estimates vary across samples:

```{r}
# take many little samples from population
samples = gss_abortion |> 
  rep_sample_n(size = 10, reps = 1000) |> 
  summarise(avg_abnomore = mean(abnomore, na.rm = TRUE))

ggplot(samples, aes(x = avg_abnomore)) + 
  geom_histogram()
```


## Bias the polls


What's the true population parameter:

```{r}
gss_abortion |> 
  summarize(abany = mean(abany, na.rm = TRUE))
```


Bias the polls:


```{r}
bias_sample = gss_abortion |> 
  filter(educ <= 5) |> 
  rep_sample_n(size = 100, reps = 5000) |> 
  summarize(abany = mean(abany, na.rm = TRUE))


ggplot(bias_sample, aes(x = abany)) + 
  geom_histogram() +
  geom_vline(xintercept = .414, size = 2, color = "red")
```



## Bootstrapping


Look at the data:

```{r}
## death penalty
issues = read_csv("https://www.dropbox.com/s/x5xncajqsz0q09l/voter-files-issues.csv?dl=1")
```


Average support for death penalty in Kansas:

```{r}
just_kansas = issues |> 
  filter(state == "Kansas")
  
just_kansas |> summarise(avg_death = mean(deathpen_2016, na.rm = TRUE))
```


Bootstrap estimates:


```{r}
boot_kansas = just_kansas |> 
  rep_sample_n(size = nrow(just_kansas), replace = TRUE,
               reps = 1000) |> 
  summarise(avg_death = mean(deathpen_2016, na.rm = TRUE))
```


Plot the estimates:

```{r}
ggplot(boot_kansas, aes(x = avg_death)) + 
  geom_histogram()
```


Calculate the standard error:


```{r}
# standard error
boot_kansas |> 
  summarise(standard_error = sd(avg_death))
```


Calculate the 95% CI:


```{r}
# 95% CI
boot_kansas |> 
  summarise(low = quantile(avg_death, probs = .025),
            high = quantile(avg_death, probs = .975))
```



