---
title: "Hypothesis testing"
---



```{r setup, include=FALSE}
# set knit options
knitr::opts_chunk$set(
  fig.width=9, 
  fig.height=5, 
  fig.retina=3,
  fig.align="center",
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

# source in functions
source(here::here("R", "funcs.R"))

```




# In-class example

Load the data and libraries:

```{r}
library(tidyverse)
library(juanr)
library(moderndive)
library(huxtable)

# data
head(wealth)
```


What percent of households in Honduras have a TV?

```{r}
wealth |> 
  summarise(r1 = mean(r1, na.rm = TRUE))
```


82.4% of households in our sample have a TV in their home. This is an *estimate* based on *one sample*. How uncertain should we feel about this estimate given our sample size and variability? 

We can bootstrap 1,000 samples with replacement of the same size as ours:

```{r}
set.seed(1990)
# how uncertain should we be?
boots = wealth |> 
  rep_sample_n(size = 1560, reps = 1000, 
               replace = TRUE) |> 
  summarise(r1_avg = mean(r1, na.rm = TRUE))
```


We can then look at the distribution of likely averages:

```{r}

ggplot(boots, aes(x = r1_avg)) + 
  geom_density() + 
  labs(x = "Bootstrapped proportion of households with TV")
```


We can quantify uncertainty using the standard error:

```{r}
## standard error?
boots |> 
  summarise(r1_avg_avg = mean(r1_avg, na.rm = TRUE),
            r1_se = sd(r1_avg, na.rm = TRUE))
```

So our best guess is $Estimate \pm 2 \times StandardError$ = $.82 + 2 \times .01 = .84, .82 - 2 \times .01 = .80$

Our best guess for percent of households in Honduras with a TV is between 80% and 84%. 


We can also look at the 95% CI:

```{r}
## confidence interval?
boots %>% 
  summarise(low = quantile(r1_avg, .025),
            mean = mean(r1_avg), 
            high = quantile(r1_avg, .975))
```


When we do regression with `lm`, R returns two measures of uncertainty: (1) the standard error, and (2) whether the 95% confidence interval excludes zero (ie., *statistical significance*)


```{r}
mod = lm(mpg ~ wt, data = mtcars)
huxreg(mod, statistics = "nobs")
```


The standard error is the number in parentheses below each coefficient. The standard error for *wt* is (.56). 

The stars tell you whether a result is statistically significant or not, and what size of confidence interval *excludes* zero. The estimate for *wt* is statistically significant, and the 99.9% (100 - .001) confidence interval. 


