---
title: "Syllabus"
---


::: {.grid .course-details .course-details-smaller}

::: {.g-col-12 .g-col-sm-6 .g-col-md-4}
### Instructor

- {{< fa user >}} &nbsp; [{{< var instructor.name >}}]({{< var instructor.url >}})
- {{< fa university >}} &nbsp; {{< var instructor.office >}}
- {{< fa envelope >}} &nbsp; <a href="mailto:{{< var instructor.email >}}">{{< var instructor.email >}}</a>
- {{< fa calendar-check >}} &nbsp; [Schedule an appointment]({{< var instructor.appointment_url >}})
:::


::: {.g-col-12 .g-col-md-4 .contact-policy}
### Teaching assistants

- {{< fa user-graduate >}}  &nbsp; Haley Daarstad
- {{< fa user-graduate >}}  &nbsp; RyuGyung ("Rio") Park
- {{< fa user-graduate >}}  &nbsp; Jack Rametta
:::

::: {.g-col-12 .g-col-sm-6 .g-col-md-4}
### Course details

- {{< fa calendar >}} &nbsp; {{< var course.days >}}
- {{< fa calendar-alt >}} &nbsp; {{< var course.dates >}}
- {{< fa clock >}} &nbsp; {{< var course.time >}}
- {{< fa location-dot >}} &nbsp; {{< var course.location >}}
:::


:::


```{r load-libraries, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tibble)
library(pander)
```




## What is this class about? 


Policymakers, academics, journalists, firms, NGOs and many others use quantitative data every day to make decisions. Data are also being used to make **causal** claims about the world -- to argue that some policy will improve or worsen our lives. 


This undergraduate course is about how to **do** data analysis and how to **think** causally about the political world. We will cover topics in data science, programming, regression, causal inference, and uncertainty. 


By the end of this course, you will be able to:


1. Feel comfortable manipulating data in R
2. Craft effective visualizations of patterns in data
3. Draw causal diagrams and identify obstacles to causal claims
4. Understand the basics of regression and uncertainty




## What materials do I need for this course?


All materials for this course are **free and online**. 



## Course Policies


### How can I get help or get in contact with the instructors?



*If you need one-on-one help you can also reach out to the TAs* to schedule a time to meet. The TAs and their office hours are at the top of this page.



*If you would like to speak with me about something that only pertains to you* (e.g., your grades, academic advice), you can sign up for office hours on [Calendly](https://calendly.com/juan-f-tellez). If there's a time-sensitive issue you can email me.


### Honor Code

Be nice. Don't cheat. The [Code of Academic Conduct](https://ossja.ucdavis.edu/code-academic-conduct) is in effect in this class and all others at the University. I will treat violations seriously. If you have doubts, it is your responsibility to ask about the Code’s application.


### Counseling & Psychiatry Services


Life at Davis can be complicated and challenging. You might feel overwhelmed, experience anxiety or depression, or struggle with relationships or family responsibilities. [UC Davis Counseling Services](https://shcs.ucdavis.edu/services/counseling-services) provide *confidential* support for students who are struggling with mental health and emotional challenges. Please do not hesitate to contact them for assistance—getting help is a smart and good thing to do.



## Assignments and grades

You can find descriptions for all the assignments on the [assignments page](/assignment/).


```{r assignments-grades, include=FALSE}
assignments <- tribble(
  ~Percent, ~Assignment,
  .15, "Weekly check-in",
  .5, "Problem sets (8)",
  .35, "Final project"
) %>%
  mutate(Percent = scales::percent(Percent))

grading <- tribble(
  ~Grade, ~Range, ~Grade1, ~Range1,
  "A",  "93–100%", "C",  "73–76%",
  "A−", "90–92%", "C−", "70–72%",
  "B+", "87–89%", "D+", "67–69%",
  "B",  "83–86%", "D",  "63–66%",
  "B−", "80–82%", "D−", "60–62%",
  "C+", "77–79%", "F",  "< 60%"
)
```

```{r show-assignments-table, echo=FALSE, results="asis"}
assignments_total <- tibble(Assignment = "Total",
                                Percent = scales::percent(1))

bind_rows(assignments, assignments_total) %>% 
  select(Assignment, Percent) %>% 
  pandoc.table(justify = "lc", missing = "—")
```


```{r show-grades-table, echo=FALSE, results="asis"}
grading %>%
  pandoc.table(justify = "lclc", 
               col.names = c("Grade", "Range", "Grade", "Range"))
```



## Old tech

Once you have read this entire syllabus and [the assignments page](/assignment/), please post a picture or gif of an old computer, something with a [cyberpunk](https://en.wikipedia.org/wiki/Cyberpunk) aesthetic, or some old technology to the Slack (I'll round your final grade up to the nearest whole number; you’ve got until September 29th).






## Credits

This course draws on code, content, ideas, inspirations and much more from work by Andrew Heiss, Nick C. Huntington-Klein, Kieran Healy, Scott Cunningham, and others who have made their courses publicly available.
