---
title: "DAGs"
---


```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details = tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## Optional

- work through the examples in the `ggdag` package vignette [ here](https://cran.r-project.org/web/packages/ggdag/vignettes/intro-to-dags.html).^[Don't just copy/paste the code; write it out yourself! Change the variables in the smoking/cancer example to something of interest to you.]
- {{< fa book >}} [Directed Acyclical Graphs in Causal Inference: The Mixtape](https://mixtape.scunning.com)^[This reading is mathy-er than what we've covered so far. It also features code from STATA, which we don't use. Don't worry about that so much; we will focus on broad concepts/ideas.]


## Slides



```{r show-slide-tabs, echo=FALSE, results="asis", eval = FALSE}
slide_buttons("/static/slides/02-viz")
slide_tabs(slide_details, "/static/slides/02-viz.html")
```

::: {.callout-note}
You can navigate through the slides with <kbd>←</kbd> and <kbd>→</kbd>. If you type <kbd>F</kbd> you can go full-screen. 
:::



